### 2022-11-14 submission/manuscript.Rmd
Initial commit

---
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
csl: asm.csl
geometry: margin=1.0in
header-includes:
 - \usepackage{booktabs}
 - \usepackage{longtable}
 - \usepackage{array}
 - \usepackage{multirow}
 - \usepackage{wrapfig}
 - \usepackage{float}
 - \usepackage{colortbl}
 - \usepackage{pdflscape}
 - \usepackage{tabu}
 - \usepackage{threeparttable}
 - \usepackage{threeparttablex}
 - \usepackage[normalem]{ulem}
 - \usepackage{makecell}
 - \usepackage{setspace}
 - \doublespacing
 - \usepackage[left]{lineno}
 - \usepackage{helvet} % Helvetica font
 - \renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
 - \usepackage[T1]{fontenc}
---

```{r knitr_settings, tidy=TRUE, eval=TRUE, echo=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("message" = FALSE)
opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
  print(x)
  if(is.list(x)){
    x <- unlist(x)
  }
  if(is.numeric(x)){
    if(abs(x - round(x)) < .Machine$double.eps^0.5){
      paste(format(x,big.mark=',', digits=0, scientific=FALSE))
    } else {
      paste(format(x,big.mark=',', digits=2, nsmall=2, scientific=FALSE))
    }
  } else {
      paste(x)
  }
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace*{10mm}

# The *Riffomonas* YouTube Channel: An Educational Resource to Foster Reproducible Research Practices

\vspace{15mm}

Running title: *Riffomonas* YouTube Channel

\vspace{15mm}

Patrick D. Schloss${^1}$${^\dagger}$

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: \href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI 48109

\vspace{35mm}

### Educational resource

\newpage
\linenumbers

## Abstract
Limit the abstract to 50 words or fewer

\newpage

Limit the paper to 500 words or fewer

[@Schloss2018a]
[@Schloss2018b]

Code Club

https://www.youtube.com/c/RiffomonasProject

Need for all scientists to strengthen their data science skills
* Focus on reproducible research methods
* command line, R, Rmarkdown, version control, data visualization, project organization

Previously...
* Developed Riffomonas reproducible research tutorial series
* Developed other online tutorials teaching scientists to use in R with microbiome (minimalR) and more general (generalR) data

Current state of the art...
* Workshop-based tutorials - intensive learning opportunities (minimalR/generalR and Carpentries)
* Books - use toy datasets, highlight individual commands

The *Riffomonas* YouTube Channel...
* Repository for Reproducible Research Tutorial series
* Code Club series

Code Club
* Born out of pandemic and need for community and desire to fill the hole in offerings for repeated practice, applying concepts in different contexts, and intgration of concepts in project-based approach
* So far...
  - Live coded the process of developing, writing, and publishing a paper
  - Use "real" and not toy datasets - microbiome, weather, commodity prices, COVID-19 vaccine attitudes - to develop concepts
* Posted 1-3 videos a week
* Recently started a free weekly email newsletter with practice problems that parallel content in the videos

Availability of code through Jekyll powered blog and repositories within a GitHub-based project

Pedagogy...
* Offering encouragement
* Normalize failure
* Repeated practice opportunities
* Ability to apply concepts in different contexts
* Integration of concepts in a project-based approach

## Acknowledgements

\newpage

## References

<div id="refs"></div>

\newpage

**Figure 1. Lorem ipsum dolor sit amet.** Consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

### 2022-11-14 submission/manuscript.tex
Initial commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\vspace*{10mm}

\hypertarget{the-riffomonas-youtube-channel-an-educational-resource-to-foster-reproducible-research-practices}{%
\section{\texorpdfstring{The \emph{Riffomonas} YouTube Channel: An
Educational Resource to Foster Reproducible Research
Practices}{The Riffomonas YouTube Channel: An Educational Resource to Foster Reproducible Research Practices}}\label{the-riffomonas-youtube-channel-an-educational-resource-to-foster-reproducible-research-practices}}

\vspace{15mm}

Running title: \emph{Riffomonas} YouTube Channel

\vspace{15mm}

Patrick D. Schloss\({^1}\)\({^\dagger}\)

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

\(1\) Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

\vspace{35mm}

\hypertarget{educational-resource}{%
\subsubsection{Educational resource}\label{educational-resource}}

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Limit the abstract to 50 words or fewer

\newpage

Limit the paper to 500 words or fewer

(1) (2)

Need for all scientists to strengthen their data science skills * Focus
on reproducible research methods * command line, R, Rmarkdown, version
control, data visualization, project organization

Previously\ldots{} * Developed Riffomonas reproducible research tutorial
series * Developed other online tutorials teaching scientists to use in
R with microbiome (minimalR) and more general (generalR) data

Current state of the art\ldots{} * Workshop-based tutorials - intensive
learning opportunities (minimalR/generalR and Carpentries) * Books - use
toy datasets, highlight individual commands

The \emph{Riffomonas} YouTube Channel\ldots{} * Repository for
Reproducible Research Tutorial series * Code Club series

Code Club * Born out of pandemic and need for community and desire to
fill the hole in offerings for repeated practice, applying concepts in
different contexts, and intgration of concepts in project-based approach
* So far\ldots{} - Live coded the process of developing, writing, and
publishing a paper - Use ``real'' and not toy datasets - microbiome,
weather, commodity prices, COVID-19 vaccine attitudes - to develop
concepts * Posted 1-3 videos a week * Recently started a free weekly
email newsletter with practice problems that parallel content in the
videos

Availability of code through Jekyll powered blog and repositories within
a GitHub-based project

Pedagogy\ldots{} * Offering encouragement * Normalize failure * Repeated
practice opportunities * Ability to apply concepts in different contexts
* Integration of concepts in a project-based approach

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-Schloss2018a}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Schloss PD}. 2018. The {R}iffomonas reproducible
research tutorial series. Journal of Open Source Education
\textbf{1}:13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}.}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2018b}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Schloss PD}. 2018. Identifying and overcoming
threats to reproducibility, replicability, robustness, and
generalizability in microbiome research. {mBio} \textbf{9}.
doi:\href{https://doi.org/10.1128/mbio.00525-18}{10.1128/mbio.00525-18}.}

\end{CSLReferences}

\newpage

\textbf{Figure 1. Lorem ipsum dolor sit amet.} Consectetur adipisicing
elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi
ut aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa
qui officia deserunt mollit anim id est laborum.

\end{document}

### 2022-12-12 submission/manuscript.Rmd
First full draft

@@
 ## Abstract
-Limit the abstract to 50 words or fewer
+
+Methods for analyzing data in a reproducible manner are often viewed as impenetrable to scientists more familiar with laboratory research.
+The *Riffomonas* YouTube channel is committed to teaching these scientists and others how to engage in reproducible research using modern data science tools.
 
@@
 
-Limit the paper to 500 words or fewer
+As high throughput data generation becomes more common in microbiology and other disciplines there is a significant need for laboratory scientists to develop data science skills [@Barone2017].
+Unfortunately, traditional undergraduate and graduate biology training programs are often deficient in opportunities for scientists to develop the skills necessary to analyze large datasets in a reproducible and robust manner [@Schloss2018b; @Williams2019].
+Numerous organizations seek to fill this void including the Carpentries, Codeacademy, and DataCamp [@Wilson2016].
+There are also numerous video tutorials available on YouTube.
+Although the content available through these platforms are popular, there has been a gap in content that emphasizes project-based learning.
 
-[@Schloss2018a]
-[@Schloss2018b]
+The *Riffomonas* YouTube channel (https://www.youtube.com/c/RiffomonasProject) seeks to fill this gap.
+I started consistently posting videos at the beginning of the COVID-19 pandemic in the Spring of 2020.
+As of the end of November 2022, the channel included 285 videos that had been viewed 635,947 times; the channel had 11,327 subscribers.
+The majority of these are 264 videos in the "Code Club" playlist [@Hagan2020].
+Other videos are related to a previously described tutorial series on reproducible research [@Schloss2018a] and series where reproducible reseach practices are used to address topical questions.
+Code Club videos are typically between 20 and 30 minutes long.
+The code that is developed in the videos is available through a website (https://riffomonas.org/code_club/) and the channel's GitHub-hosted account (https://github.com/riffomonas).
 
-Code Club
+The name, *Riffomonas*, comes from the concept of "riffing" where musical themes are adapted to achieve a similar sound, albeit perhaps in a different context [@Schloss2018a].
+This is to emphasize the value of reproducibility not only to recreate a set of results but to apply a method with a different dataset [@Leek2015].
+The channel covers topics related to reproducible data analysis practices including R programming, data visualization, project organization, version control, command line programming, workflow tools, and scientific publishing.
+Each video includes a brief introduction followed by me live coding to achieve a goal.
+I emphasize the use of live coding to modulate the rate of instruction and to show viewers my own coding practices.
+Observing a experienced analyst make mistakes normalizes some level of failure and demonstrates the strategies they can use to resolve their own mistakes.
+Viewers are encouraged to follow along with each video and to apply the new information to their own project.
 
-https://www.youtube.com/c/RiffomonasProject
-
-Need for all scientists to strengthen their data science skills
-* Focus on reproducible research methods
-* command line, R, Rmarkdown, version control, data visualization, project organization
-
-Previously...
-* Developed Riffomonas reproducible research tutorial series
-* Developed other online tutorials teaching scientists to use in R with microbiome (minimalR) and more general (generalR) data
-
-Current state of the art...
-* Workshop-based tutorials - intensive learning opportunities (minimalR/generalR and Carpentries)
-* Books - use toy datasets, highlight individual commands
-
-The *Riffomonas* YouTube Channel...
-* Repository for Reproducible Research Tutorial series
-* Code Club series
-
-Code Club
-* Born out of pandemic and need for community and desire to fill the hole in offerings for repeated practice, applying concepts in different contexts, and intgration of concepts in project-based approach
-* So far...
-  - Live coded the process of developing, writing, and publishing a paper
-  - Use "real" and not toy datasets - microbiome, weather, commodity prices, COVID-19 vaccine attitudes - to develop concepts
-* Posted 1-3 videos a week
-* Recently started a free weekly email newsletter with practice problems that parallel content in the videos
-
-Availability of code through Jekyll powered blog and repositories within a GitHub-based project
-
-Pedagogy...
-* Offering encouragement
-* Normalize failure
-* Repeated practice opportunities
-* Ability to apply concepts in different contexts
-* Integration of concepts in a project-based approach
+Each video emphasizes a specific topic, but includes other content that is selected to review topics covered in recent videos.
+Although videos can be watched individually, they often form a project arc.
+For example, between July 2020 and July 2021, I formulated a research question, obtained and analyzed data to answer the question and wrote a paper that was published in *mSphere* [@Schloss2021].
+This series of 67 videos covered every topic from creating the initial directory on my computer to house the project files through reviewing the proofs of the published manuscript.
+Other project arcs have included visualizing microbiome data, modeling microbiome data using machine learning tools, analyzing the impacts of rarefying microbiome data, and other topics.
+Going forward, the *Riffomonas* channel will continue to post project-based content to help researchers develop their reproducible research skills.
 
 ## Acknowledgements
+
+I am grateful to the audience of the *Riffomonas* channel for their feedback on topics that I should cover in future episodes.
 
@@
 <div id="refs"></div>
-
-\newpage
-
-**Figure 1.
-Lorem ipsum dolor sit amet.** Consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

### 2022-12-12 submission/manuscript.tex
Make sure plumbing works properly
First full draft

@@
 
-Limit the abstract to 50 words or fewer
+Methods for analyzing data in a reproducible manner are often viewed as
+impenetrable to scientists more familiar with laboratory research. The
+\emph{Riffomonas} YouTube channel is committed to teaching these
+scientists and others how to engage in reproducible research using
+modern data science tools.
 
@@
 
-Limit the paper to 500 words or fewer
+As high throughput data generation becomes more common in microbiology
+and other disciplines there is a significant need for laboratory
+scientists to develop data science skills (1). Unfortunately,
+traditional undergraduate and graduate biology training programs are
+often deficient in opportunities for scientists to develop the skills
+necessary to analyze large datasets in a reproducible and robust manner
+(2, 3). Numerous organizations seek to fill this void including the
+Carpentries, Codeacademy, and DataCamp (4). There are also numerous
+video tutorials available on YouTube. Although the content available
+through these platforms are popular, there has been a gap in content
+that emphasizes project-based learning.
 
-(1) (2)
+The \emph{Riffomonas} YouTube channel
+(\url{https://www.youtube.com/c/RiffomonasProject}) seeks to fill this
+gap. I started consistently posting videos at the beginning of the
+COVID-19 pandemic in the Spring of 2020. As of the end of November 2022,
+the channel included 285 videos that had been viewed 635,947 times; the
+channel had 11,327 subscribers. The majority of these are 264 videos in
+the ``Code Club'' playlist (5). Other videos are related to a previously
+described tutorial series on reproducible research (6) and series where
+reproducible reseach practices are used to address topical questions.
+Code Club videos are typically between 20 and 30 minutes long. The code
+that is developed in the videos is available through a website
+(\url{https://riffomonas.org/code_club/}) and the channel's
+GitHub-hosted account (\url{https://github.com/riffomonas}).
 
-Need for all scientists to strengthen their data science skills * Focus
-on reproducible research methods * command line, R, Rmarkdown, version
-control, data visualization, project organization
+The name, \emph{Riffomonas}, comes from the concept of ``riffing'' where
+musical themes are adapted to achieve a similar sound, albeit perhaps in
+a different context (6). This is to emphasize the value of
+reproducibility not only to recreate a set of results but to apply a
+method with a different dataset (7). The channel covers topics related
+to reproducible data analysis practices including R programming, data
+visualization, project organization, version control, command line
+programming, workflow tools, and scientific publishing. Each video
+includes a brief introduction followed by me live coding to achieve a
+goal. I emphasize the use of live coding to modulate the rate of
+instruction and to show viewers my own coding practices. Observing a
+experienced analyst make mistakes normalizes some level of failure and
+demonstrates the strategies they can use to resolve their own mistakes.
+Viewers are encouraged to follow along with each video and to apply the
+new information to their own project.
 
-Previously\ldots{} * Developed Riffomonas reproducible research tutorial
-series * Developed other online tutorials teaching scientists to use in
-R with microbiome (minimalR) and more general (generalR) data
-
-Current state of the art\ldots{} * Workshop-based tutorials - intensive
-learning opportunities (minimalR/generalR and Carpentries) * Books - use
-toy datasets, highlight individual commands
-
-The \emph{Riffomonas} YouTube Channel\ldots{} * Repository for
-Reproducible Research Tutorial series * Code Club series
-
-Code Club * Born out of pandemic and need for community and desire to
-fill the hole in offerings for repeated practice, applying concepts in
-different contexts, and intgration of concepts in project-based approach
-* So far\ldots{} - Live coded the process of developing, writing, and
-publishing a paper - Use ``real'' and not toy datasets - microbiome,
-weather, commodity prices, COVID-19 vaccine attitudes - to develop
-concepts * Posted 1-3 videos a week * Recently started a free weekly
-email newsletter with practice problems that parallel content in the
-videos
-
-Availability of code through Jekyll powered blog and repositories within
-a GitHub-based project
-
-Pedagogy\ldots{} * Offering encouragement * Normalize failure * Repeated
-practice opportunities * Ability to apply concepts in different contexts
-* Integration of concepts in a project-based approach
+Each video emphasizes a specific topic, but includes other content that
+is selected to review topics covered in recent videos. Although videos
+can be watched individually, they often form a project arc. For example,
+between July 2020 and July 2021, I formulated a research question,
+obtained and analyzed data to answer the question and wrote a paper that
+was published in \emph{mSphere} (8). This series of 67 videos covered
+every topic from creating the initial directory on my computer to house
+the project files through reviewing the proofs of the published
+manuscript. Other project arcs have included visualizing microbiome
+data, modeling microbiome data using machine learning tools, analyzing
+the impacts of rarefying microbiome data, and other topics. Going
+forward, the \emph{Riffomonas} channel will continue to post
+project-based content to help researchers develop their reproducible
+research skills.
 
@@
 \subsection{Acknowledgements}\label{acknowledgements}}
+
+I am grateful to the audience of the \emph{Riffomonas} channel for their
+feedback on topics that I should cover in future episodes.
 
@@
 \begin{CSLReferences}{0}{1}
-\leavevmode\vadjust pre{\hypertarget{ref-Schloss2018a}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Barone2017}{}}%
 \CSLLeftMargin{1. }%
-\CSLRightInline{\textbf{Schloss PD}. 2018. The {R}iffomonas reproducible
-research tutorial series. Journal of Open Source Education
-\textbf{1}:13.
-doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}.}
+\CSLRightInline{\textbf{Barone L}, \textbf{Williams J}, \textbf{Micklos
+D}. 2017. Unmet needs for analyzing biological big data: A survey of 704
+{NSF} principal investigators. {PLOS} Computational Biology
+\textbf{13}:e1005755.
+doi:\href{https://doi.org/10.1371/journal.pcbi.1005755}{10.1371/journal.pcbi.1005755}.}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Williams2019}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{\textbf{Williams JJ}, \textbf{Drew JC},
+\textbf{Galindo-Gonzalez S}, \textbf{Robic S}, \textbf{Dinsdale E},
+\textbf{Morgan WR}, \textbf{Triplett EW}, \textbf{Burnette JM},
+\textbf{Donovan SS}, \textbf{Fowlks ER}, \textbf{Goodman AL},
+\textbf{Grandgenett NF}, \textbf{Goller CC}, \textbf{Hauser C},
+\textbf{Jungck JR}, \textbf{Newman JD}, \textbf{Pearson WR},
+\textbf{Ryder EF}, \textbf{Sierk M}, \textbf{Smith TM},
+\textbf{Tosado-Acevedo R}, \textbf{Tapprich W}, \textbf{Tobin TC},
+\textbf{Toro-Martínez A}, \textbf{Welch LR}, \textbf{Wilson MA},
+\textbf{Ebenbach D}, \textbf{McWilliams M}, \textbf{Rosenwald AG},
+\textbf{Pauley MA}. 2019. Barriers to integration of bioinformatics into
+undergraduate life sciences education: A national study of {US} life
+sciences faculty uncover significant barriers to integrating
+bioinformatics into undergraduate instruction. {PLOS} {ONE}
+\textbf{14}:e0224288.
+doi:\href{https://doi.org/10.1371/journal.pone.0224288}{10.1371/journal.pone.0224288}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wilson2016}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{\textbf{Wilson G}. 2016. Software carpentry: Lessons
+learned. F1000Research.
+doi:\href{https://doi.org/10.12688/f1000research.3-62.v2}{10.12688/f1000research.3-62.v2}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hagan2020}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{\textbf{Hagan AK}, \textbf{Lesniak NA}, \textbf{Balunas
+MJ}, \textbf{Bishop L}, \textbf{Close WL}, \textbf{Doherty MD},
+\textbf{Elmore AG}, \textbf{Flynn KJ}, \textbf{Hannigan GD},
+\textbf{Koumpouras CC}, \textbf{Jenior ML}, \textbf{Kozik AJ},
+\textbf{McBride K}, \textbf{Rifkin SB}, \textbf{Stough JMA},
+\textbf{Sovacool KL}, \textbf{Sze MA}, \textbf{Tomkovich S},
+\textbf{Topcuoglu BD}, \textbf{Schloss PD}. 2020. Ten simple rules to
+increase computational skills among biologists with code clubs. {PLOS}
+Computational Biology \textbf{16}:e1008119.
+doi:\href{https://doi.org/10.1371/journal.pcbi.1008119}{10.1371/journal.pcbi.1008119}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schloss2018a}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{\textbf{Schloss PD}. 2018. The {R}iffomonas reproducible
+research tutorial series. Journal of Open Source Education
+\textbf{1}:13.
+doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Leek2015}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{\textbf{Leek JT}, \textbf{Peng RD}. 2015. Reproducible
+research can still be wrong: Adopting a prevention approach. Proceedings
+of the National Academy of Sciences \textbf{112}:1645--1646.
+doi:\href{https://doi.org/10.1073/pnas.1421412111}{10.1073/pnas.1421412111}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schloss2021}{}}%
+\CSLLeftMargin{8. }%
+\CSLRightInline{\textbf{Schloss PD}. 2021. Amplicon sequence variants
+artificially split bacterial genomes into separate clusters. {mSphere}
+\textbf{6}.
+doi:\href{https://doi.org/10.1128/msphere.00191-21}{10.1128/msphere.00191-21}.}
+
 \end{CSLReferences}
 
-\newpage
-
-\textbf{Figure 1. Lorem ipsum dolor sit amet.} Consectetur adipisicing
-elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi
-ut aliquip ex ea commodo consequat. Duis aute irure dolor in
-reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
-pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa
-qui officia deserunt mollit anim id est laborum.
-
 \end{document}

### 2022-12-12 submission/manuscript.Rmd
Editing

@@
 output:
+  word_document: default
   pdf_document:
@@
  - \usepackage[left]{lineno}
+ - \usepackage[none]{hyphenat}
+ - \raggedright
  - \usepackage{helvet} % Helvetica font
@@
 Although videos can be watched individually, they often form a project arc.
-For example, between July 2020 and July 2021, I formulated a research question, obtained and analyzed data to answer the question and wrote a paper that was published in *mSphere* [@Schloss2021].
+For example, between July 2020 and July 2021, I formulated a research question, obtained and analyzed data to answer the question, and wrote a paper that was published in *mSphere* [@Schloss2021].
 This series of 67 videos covered every topic from creating the initial directory on my computer to house the project files through reviewing the proofs of the published manuscript.

### 2022-12-12 submission/manuscript.tex
Editing

@@
 \usepackage[left]{lineno}
+\usepackage[none]{hyphenat}
+\raggedright
 \usepackage{helvet} % Helvetica font
@@
 between July 2020 and July 2021, I formulated a research question,
-obtained and analyzed data to answer the question and wrote a paper that
-was published in \emph{mSphere} (8). This series of 67 videos covered
-every topic from creating the initial directory on my computer to house
-the project files through reviewing the proofs of the published
+obtained and analyzed data to answer the question, and wrote a paper
+that was published in \emph{mSphere} (8). This series of 67 videos
+covered every topic from creating the initial directory on my computer
+to house the project files through reviewing the proofs of the published
 manuscript. Other project arcs have included visualizing microbiome

### 2022-12-20 submission/manuscript.Rmd
Render with table, edits, and track changes

@@
 The *Riffomonas* YouTube channel (https://www.youtube.com/c/RiffomonasProject) seeks to fill this gap.
-I started consistently posting videos at the beginning of the COVID-19 pandemic in the Spring of 2020.
+I started consistently posting videos at the beginning of the COVID-19 pandemic in April 2020.
 As of the end of November 2022, the channel included 285 videos that had been viewed 635,947 times; the channel had 11,327 subscribers.
-The majority of these are 264 videos in the "Code Club" playlist [@Hagan2020].
+The majority of these are 264 videos in the "Code Club" playlist [@Hagan2020] (Table 1).
 Other videos are related to a previously described tutorial series on reproducible research [@Schloss2018a] and series where reproducible reseach practices are used to address topical questions.
@@
 This is to emphasize the value of reproducibility not only to recreate a set of results but to apply a method with a different dataset [@Leek2015].
-The channel covers topics related to reproducible data analysis practices including R programming, data visualization, project organization, version control, command line programming, workflow tools, and scientific publishing.
+The channel covers topics related to reproducible data analysis practices including R programming, data visualization, project organization, version control, command line programming, workflow tools, and scientific publishing (Table 1).
 Each video includes a brief introduction followed by me live coding to achieve a goal.
@@
 Each video emphasizes a specific topic, but includes other content that is selected to review topics covered in recent videos.
-Although videos can be watched individually, they often form a project arc.
+Although videos can be watched individually, they often form a project arc (Table 1).
 For example, between July 2020 and July 2021, I formulated a research question, obtained and analyzed data to answer the question, and wrote a paper that was published in *mSphere* [@Schloss2021].
@@
 <div id="refs"></div>
+
+\newpage
+
+**Table 1.
+Description of the data science topics and project based series covered in the playlists found on the *Riffomonas* YouTube Channel.** Because most videos cover more than one topic they are found in mutiple playlists.
+Playlists and counts were current as of December 1, 2022.
+Playlists can be found under the Playlist tab at https://www.youtube.com/c/RiffomonasProject.
+
+| |  **Number of videos**  |
+|:-----------------------------------------|:----------------------:|
+| **Videos covering data science topics** |  |
+| [Data visualization with R's tidyverse and allied packages](https://youtube.com/playlist?list=PLmNrK_nkqBpKHYucwmN6YlNjoyeWoGBYu) | 146 |
+| [Data manipulation within R's tidyverse and other packages](https://youtube.com/playlist?list=PLmNrK_nkqBpKf7j_ewpUm-w33R6PJYtD9) | 116 |
+| [Data analysis with base R](https://youtube.com/playlist?list=PLmNrK_nkqBpLaDgAZb5SHHhJ8TDufVv7H) | 39 |
+| [Tools for reproducible data analysis](https://youtube.com/playlist?list=PLmNrK_nkqBpLdTEOj_-bwp58490i8ezRI) | 33 |
+| [Working at the command line](https://youtube.com/playlist?list=PLmNrK_nkqBpKhP0Jb2XGPoRRfToHF1w-y) | 26 |
+| [Literate programming with R markdown](https://youtube.com/playlist?list=PLmNrK_nkqBpKT5Wz1IHZneRXPUGNtB-IG) | 18 |
+| [Machine learning with mikropml R package](https://youtube.com/playlist?list=PLmNrK_nkqBpKpzb9-vI4V7SdXC-jXEcmg) | 16 |
+| [Version control with git and GitHub](https://youtube.com/playlist?list=PLmNrK_nkqBpJtNdQBPhPWjIFRYeFOGfJ1) | 15 |
+| [Scientific writing](https://youtube.com/playlist?list=PLmNrK_nkqBpI7_TMg72ub3Nflk9F8qPlb) | 15 |
+| [Project organization](https://youtube.com/playlist?list=PLmNrK_nkqBpKtCHqSSHXZdK-C4Zg89Rrv) | 3 |
+|    |
+| **Project based series** |  |
+| [All Code Club videos since April 2, 2020](https://youtube.com/playlist?list=PLmNrK_nkqBpKY3SZiivlIGvcLX-KHmfR8) | 265 |
+| [Microbiome data analysis and visualization](https://youtube.com/playlist?list=PLmNrK_nkqBpIIRdQTS2aOs5OD7vVMKWAi) | 86 |
+| [ASV/OTU senstitivity and specificity analyses](https://youtube.com/playlist?list=PLmNrK_nkqBpL7m_tyWdQgdyurerttCsPY) | 67 |
+| [Visualizing COVID-19 vaccination attitudes](https://youtube.com/playlist?list=PLmNrK_nkqBpL9STHqSod2LGwciLk9TXZ1) | 31 |
+| [Climate change data visualization](https://youtube.com/playlist?list=PLmNrK_nkqBpJTSHf3IsN_K_pjFu58z9Oq) | 29 |
+| [Evaluating rarefaction and its alternatives](https://youtube.com/playlist?list=PLmNrK_nkqBpJuhS93PYC-Xr5oqur7IIWf) | 18 |
+| [Drought index visualization](https://youtube.com/playlist?list=PLmNrK_nkqBpK6iqwN3QeQyXqI6DrcGgIm) | 17 |
+| [Reproducible reserach tutorial series](https://youtube.com/playlist?list=PLmNrK_nkqBpL0d2E26TqPkmTAfelYKbQX) | 14 |
+| [Commemorating Juneteenth 2022 with a visualization](https://youtube.com/playlist?list=PLmNrK_nkqBpK24Xzukw7ngIDEeukkunzO) | 5 |
+| [2018 MLB All Star Break data analysis sprint](https://youtube.com/playlist?list=PLmNrK_nkqBpLd_IvW37Fm1w5LRv9ggvYI) | 4 |

### 2022-12-20 submission/manuscript.tex
Render with table, edits, and track changes

@@
 \usepackage[margin=1.0in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 gap. I started consistently posting videos at the beginning of the
-COVID-19 pandemic in the Spring of 2020. As of the end of November 2022,
-the channel included 285 videos that had been viewed 635,947 times; the
+COVID-19 pandemic in April 2020. As of the end of November 2022, the
+channel included 285 videos that had been viewed 635,947 times; the
 channel had 11,327 subscribers. The majority of these are 264 videos in
-the ``Code Club'' playlist (5). Other videos are related to a previously
-described tutorial series on reproducible research (6) and series where
-reproducible reseach practices are used to address topical questions.
-Code Club videos are typically between 20 and 30 minutes long. The code
-that is developed in the videos is available through a website
-(\url{https://riffomonas.org/code_club/}) and the channel's
+the ``Code Club'' playlist (5) (Table 1). Other videos are related to a
+previously described tutorial series on reproducible research (6) and
+series where reproducible reseach practices are used to address topical
+questions. Code Club videos are typically between 20 and 30 minutes
+long. The code that is developed in the videos is available through a
+website (\url{https://riffomonas.org/code_club/}) and the channel's
 GitHub-hosted account (\url{https://github.com/riffomonas}).
@@
 visualization, project organization, version control, command line
-programming, workflow tools, and scientific publishing. Each video
-includes a brief introduction followed by me live coding to achieve a
-goal. I emphasize the use of live coding to modulate the rate of
-instruction and to show viewers my own coding practices. Observing a
+programming, workflow tools, and scientific publishing (Table 1). Each
+video includes a brief introduction followed by me live coding to
+achieve a goal. I emphasize the use of live coding to modulate the rate
+of instruction and to show viewers my own coding practices. Observing a
 experienced analyst make mistakes normalizes some level of failure and
@@
 is selected to review topics covered in recent videos. Although videos
-can be watched individually, they often form a project arc. For example,
-between July 2020 and July 2021, I formulated a research question,
-obtained and analyzed data to answer the question, and wrote a paper
-that was published in \emph{mSphere} (8). This series of 67 videos
+can be watched individually, they often form a project arc (Table 1).
+For example, between July 2020 and July 2021, I formulated a research
+question, obtained and analyzed data to answer the question, and wrote a
+paper that was published in \emph{mSphere} (8). This series of 67 videos
 covered every topic from creating the initial directory on my computer
@@
 
+\newpage
+
+\textbf{Table 1. Description of the data science topics and project
+based series covered in the playlists found on the \emph{Riffomonas}
+YouTube Channel.} Because most videos cover more than one topic they are
+found in mutiple playlists. Playlists and counts were current as of
+December 1, 2022. Playlists can be found under the Playlist tab at
+\url{https://www.youtube.com/c/RiffomonasProject}.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6364}}
+  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3636}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\textbf{Number of videos}
+\end{minipage} \\
+\midrule()
+\endhead
+\textbf{Videos covering data science topics} & \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKHYucwmN6YlNjoyeWoGBYu}{Data
+visualization with R's tidyverse and allied packages} & 146 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKf7j_ewpUm-w33R6PJYtD9}{Data
+manipulation within R's tidyverse and other packages} & 116 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpLaDgAZb5SHHhJ8TDufVv7H}{Data
+analysis with base R} & 39 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpLdTEOj_-bwp58490i8ezRI}{Tools
+for reproducible data analysis} & 33 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKhP0Jb2XGPoRRfToHF1w-y}{Working
+at the command line} & 26 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKT5Wz1IHZneRXPUGNtB-IG}{Literate
+programming with R markdown} & 18 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKpzb9-vI4V7SdXC-jXEcmg}{Machine
+learning with mikropml R package} & 16 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpJtNdQBPhPWjIFRYeFOGfJ1}{Version
+control with git and GitHub} & 15 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpI7_TMg72ub3Nflk9F8qPlb}{Scientific
+writing} & 15 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKtCHqSSHXZdK-C4Zg89Rrv}{Project
+organization} & 3 \\
+& \\
+\textbf{Project based series} & \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpKY3SZiivlIGvcLX-KHmfR8}{All
+Code Club videos since April 2, 2020} & 265 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpIIRdQTS2aOs5OD7vVMKWAi}{Microbiome
+data analysis and visualization} & 86 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpL7m_tyWdQgdyurerttCsPY}{ASV/OTU
+senstitivity and specificity analyses} & 67 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpL9STHqSod2LGwciLk9TXZ1}{Visualizing
+COVID-19 vaccination attitudes} & 31 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpJTSHf3IsN_K_pjFu58z9Oq}{Climate
+change data visualization} & 29 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpJuhS93PYC-Xr5oqur7IIWf}{Evaluating
+rarefaction and its alternatives} & 18 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpK6iqwN3QeQyXqI6DrcGgIm}{Drought
+index visualization} & 17 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpL0d2E26TqPkmTAfelYKbQX}{Reproducible
+reserach tutorial series} & 14 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpK24Xzukw7ngIDEeukkunzO}{Commemorating
+Juneteenth 2022 with a visualization} & 5 \\
+\href{https://youtube.com/playlist?list=PLmNrK_nkqBpLd_IvW37Fm1w5LRv9ggvYI}{2018
+MLB All Star Break data analysis sprint} & 4 \\
+\bottomrule()
+\end{longtable}
+
 \end{document}

### 2022-12-21 submission/response_to_reviewers.md
Add response to reviewers document

---
geometry: margin=1.0in
fontfamily: helvet
fontsize: 14pt
header-includes:
  - \raggedright
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage[T1]{fontenc}
---

**Reviewer #1:**

**This manuscript describes an educational resource - the Riffomonas YouTube channel, developed by the author. I am familiar with the channel and some of its content - in order to make this a more compelling MRA, better descriptions of the resource would be nice. For example, would you be able to provide a table links to specific episodes or modules and what's covered in that module? Would you be able to use this table to also guide readers as to what they could do with that content they learn?**

**Reviewer #2:**

**First, this is an awesome resource, so I'm glad the author is choosing to report it in this venue.**

**Second, there needs to be more information on the specific content. A table is a good idea. There are playlists that could be referenced, but a more detailed description of what someone could hope to gain from going to the site would make this a more valuable resource since this could serve as the entry point for new users. E.g., what kind of code is taught (R, Python, etc.)? What kind of data types are analyzed (16S amplicon, metagenomics, etc.)? What kind of specific skills can be learned (microbiome data analysis, database curation, working with git, multivariate statistics, etc.)?**

I appreciate the enthusiasm that the reviewers both shared for the manuscript that I submitted to describe the *Riffomonas* YouTube channel as an educational resource. I have added a table that lists the various playlists and the number of videos in each that that are available on the channel. I have given the playlists titles that I hope are sufficiently descriptive to indicate they type of topics they contain. The new table includes a section indicating the general topics covered in the videos and a section that lists the types of projects that I developed over the videos. As I will continue to add videos to the channel in 2023 and beyond, I think the broad overview of topics and projects presented in Table 1 serves as an illustrative description of the videos. Because a single video may cover multiple topics the numbers do not add up to the total number of videos.

The submitted version of the manuscript embeds hyperlinks in each of the titles of the playlists. I realize that this format may not work for a printed PDF, but am optimistic that it would be possible to retain the hyperlinked titles for the online HTML and PDF versions. Unfortunately, the playlist hyperlinks are really not well formatted for printing. For example, the videos in the "Data visualization with R's tidyverse and allied packages" playlist can be found at https://youtube.com/playlist?list=PLmNrK_nkqBpKHYucwmN6YlNjoyeWoGBYu
